# pre-class


Make sure you commit this often with meaningful messages. 




### Question 1:

Using a loop, print the integers from 1 to 50. 
```{r}
for(i in 1:50){
 {
    print(i)
  }
}
```
### Question 2:

A.  Using a loop, add all the integers between 0 and 1000.
```{r}
sum=0
for(i in 1:1000){
 {
    sum=sum+i
  }
}
sum
```
B. Now, add all the EVEN integers between 0 and 1000 (hint: use seq())
```{r}
sum=0
for(i in 1:1000){
 {
   if(i %% 2==0){
    sum=sum+i
   }
  }
}
sum
```
C. Now, repeat A and B WITHOUT using a loop.
```{r}
## for A
a=seq(1:1000)
sum_a=sum(a)
sum_a
## for B
b=seq(0,1000,by=2)
sum_b=sum(b)
sum_b
```

### Question 3:

Here is a dataframe of survey data containing 5 questions :

```{r, eval=FALSE}
survey <- data.frame(
                     "participant" = c(1, 2, 3, 4, 5, 6),
                     "q1" = c(5, 3, 2, 7, 11, 0),
                     "q2" = c(4, 2, 2, 5, -10, 99),
                     "q3" = c(-4, -3, 4, 2, 9, 10),
                     "q4" = c(-30, 5, 2, 23, 4, 2),
                     "q5" = c(88, 4, -20, 2, 4, 2)
                     )
```
The response to each question should be an integer between 1 and 5. Obviously, we have some bad values in the dataframe. The goal of this problem is to fix them.

A. Using a loop, create a new dataframe called survey.clean where all the invalid values (those that are not integers between 1 and 5) are set to NA.
```{r}
survey.clean = survey
for(i in 1:dim(survey)[1]){
  for(j in 2:dim(survey)[2]){
    if(survey[i,j]>=1 & survey[i,j]<=5){
      survey.clean[i,j]=survey[i,j]} else{
        survey.clean[i,j]=NA
      }
    } 
}
survey.clean
```
B. Now, again using a loop, add a new column to the dataframe called “invalid.answers” that indicates, for each participant, how many bad answers they gave.
```{r}
invalid.answers=vector(length=dim(survey)[1],mode="numeric")
survey.new=cbind(survey.clean,invalid.answers)
survey.new=as.data.frame(survey.new)

for(i in 1:dim(survey.new)[1]){
  for(j in 2:dim(survey.new)[2]-1){
    if(is.na(survey.new[i,j])){
    survey.new[i,dim(survey.new)[2]]=survey.new[i,dim(survey.new)[2]]+1  
    }
  }
}
survey.new
```