# pre-class


Make sure you commit this often with meaningful messages. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

```{r}
##To check whether the second string is identical to the beiginning characters in the first string which have the same length of the the second  string.
check_identical <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}
## Delete the last element in a vector if the lenfth of the vector is greater than or equal to 2.
Delete_last <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}
## Repeat variable y k times where k is equal to the length of variable x.
repeat_variable <- function(x, y) {
  rep(y, length.out = length(x))
}
```

2. Compare and contrast rnorm() and MASS::mvrnorm(). How could you make them more consistent?
```{r}
library("MASS")
##The rnorm returns a vector of one-dimention normal distribution data which correspond to the the mean and variance while the mvnorm returns a matrix of high-dimention normal distribution data which correspond to the means and corvariance matrix
mvr_to_r=function(n,mu,Sigma){
  stopifnot(length(mu)==1)
  as.vector(mvrnorm(n,mu,Sigma))
}
mvr_to_r(10,15)
## We give the limitation so that the mvrnorm is only able to work with 1-dimention data
```

3. Use `lapply()` and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset. 
```{r}
data(mtcars)
coef=lapply(mtcars, function(x,na.rm=true) sd(x)/mean(x))
coef
```
4. Use vapply() to:
    a. Compute the standard deviation of every column in a numeric data frame.
    b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to use vapply() twice.)
```{r}
##a
sd_vapply_numeric=function(df){
vapply(df, function(x,na.rm=true) sd(x),numeric(1))
}
sd_vapply_numeric(mtcars)
##b
sd_vapply_mix=function(df){
  df=df[vapply(df,is.numeric,logical(1))]
  vapply(df, function(x,na.rm=true) sd(x),numeric(1))
}
b=data.frame(x=c(1,2,3,4,5),y=c(2:6),z=c(3:7),t=c("a","b","c","d","e"))
sd_vapply_mix(b)
```